### 2020-10-20 Energy_8000/highligths.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\begin{document}
%\section{Highlights (old)}
%\begin{enumerate}
%    \item A two-stage stochastic mixed complementarity capacity investment model with CO2 trading markets.

 %   \item CO2 markets for allowance allocation and allowance re-trading with endogenous price determination.

%\item Current Chilean GHG intensity pledge does not drive an important shift in the future Chilean electric mix.

%\item A more stringent carbon budget can result in carbon prices approximately ten times higher than the current price.

%\item Achieving emissions reduction goal does not necessarily results in phasing out coal in the longer term.

%\end{enumerate}
\section{Revised Highlights}

\begin{enumerate}
    \item A two-stage stochastic capacity investment equilibrium model with CO2 markets.

    \item CO2 markets for allowance allocation and re-trade with endogenous prices.

    \item Current Chilean GHG intensity pledge does not drive a shift in the generation mix.

    \item Effective carbon prices should be at least ten times larger than the Chilean tax.

    \item Strategies to comply with the renewable energy targets by mid-century.
\end{enumerate}
\end{document}