### 2021-10-04 docs/Informe/example331.tex
some order

\documentclass[12pt, letterpaper]{article}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[gen]{eurosym}
\usepackage{amsmath}
\usepackage{amssymb}

\title{Example}
\author{Amigo, Cea, Feijoo}
\date{April 2019}

\begin{document}
\maketitle

Example 3.3.1 from paper de Maere et. al. 2018 (working paper)

\section{The problem}
In this example, there is one producer that can invest in one technology. 
The plant annual capital expenditure is $I= 90 \ \euro{} / kW$ and an operating cost of $60 \ \euro{} / MWh$.
The year is represented in a single segment of $\tau=8760$ hours. The producer is risk neutral.

The equation to minimize the producer cost of operation and initial investment is 
\[ \min_{x,Y}  Ix + \tau \mathbb{E}_{\Theta} [(C-P_{\omega}) Y_{\omega}]  \textrm{\  subject to \ }  0 \leq Y_{\omega} \leq x \]

The consumer is a price-taker agent with  a quadratic utility function $U_\omega(Q_\omega) = A_\omega Q_\omega - \frac{B}{2} Q_\omega^2$ and solves
\[ \min_{Q} \tau \mathbb{E}_\Theta \left[P_\omega Q_\omega - A_\omega Q_\omega + \frac{B}{2} Q_\omega^2 \right]\]

The quadratic term B is constant across the scenarios, with $B=1 \euro{}/MWh^2$.
The linear term $A_\omega$ is the only random parameter in the example. There are 5 scenarios ($K=5 \ , \ \mathcal{Z}=\mathbb{R}^5$), and an equal physical probability of $\Theta_\omega = 0.2$ for all $\omega$. The value of A is

\begin{center}
\begin{tabular}{ c | c c c c c }
 $\omega$             & scen1 & scen2 & scen3 & scen4 & scen5 \\ 
 \hline
  $A \ [ \euro{}/MWh ] $ & 300   & 350   & 400   & 450   & 500 \\  
\end{tabular}
\end{center}

\section{Mixed complementarity problem}

\subsection{The producer}
The Lagrangian function for the producer problem is ($\lambda_1$ and $\lambda_2$ are Lagrange multipliers)

\[ \mathcal{L}= Ix + \tau \mathbb{E}_\Theta [(C-P_\omega) Y_\omega ] + \sum_w \lambda_{1,w} (Y_\omega -x) + \sum_w \lambda_{2,\omega}(0-Y_\omega) \]

The Karush-Kuhn-Tucker conditions  are:

\[ \frac{\partial \mathcal{L}}{\partial x}= I- \sum_{\omega}\lambda_{1,w} = 0  \]

\[\frac{\partial \mathcal{L}}{\partial Y_{\omega}} = \tau \Theta_\omega ( C - P_\omega ) + \lambda_{1, \omega} - \lambda_{2,\omega} =  0,\quad\omega\in\Omega \]

\vspace{1cm}

Primal feasibility :
\[ Y_\omega - x \leq 0 \quad  \forall \omega  \]
\[ 0 \leq Y_\omega \quad \forall \omega \]

Complementarity slackness:

\[ (Y_\omega - x) \cdot \lambda_{1,\omega} = 0 \]
\[ Y_\omega \cdot \lambda_{2,\omega} = 0\]

Dual feasibility

\[ \lambda_{1,\omega} \geq 0 \quad \forall \omega \]
\[ \lambda_{2,\omega} \geq 0 \quad \forall \omega \]

We can write the KKT conditions as MCP as follows

\[ 0 \leq I- \sum_{\omega}\lambda_{1,w} \perp x \geq 0 \]

\[ 0 \leq \tau \Theta_\omega C - \tau \Theta_\omega P_\omega + \lambda_{1,\omega} - \lambda_{2,\omega} \perp Y_\omega \geq 0, \quad \forall \omega\]

\[ 0 \leq x - Y_\omega \perp \lambda_{1,\omega} \geq 0 , \quad \forall \omega \]
\[ 0 \leq Y_\omega \perp \lambda_{2,\omega} \geq 0, \quad \forall \omega \]

\subsection{The consumer}

The KKT conditions for the consumer are as follows:

\[ 0 \leq \tau \Theta (P_\omega - A_\omega + B Q_\omega) \perp Q_\omega \geq 0\]

\subsection{Market clearing conditions}

 The market clearing constraint of the problem is 
\[ Y_\omega = Q_\omega \quad (P_\omega \  \textrm{dual})\]

\end{document}