### 2021-10-04 docs/DocumentoMemoria/chapters/chapter01.tex

% cap1.tex

\chapter{Introducción}
\label{c1} % la etiqueta para referencias

TEXTO