### 2022-05-20 code/Julia.Rmd
Julia Rmd with Gurobi

---
title: "Julia and Gurobi in RMarkdown"
author: "Changcheng Li"
date: "`r Sys.Date()`"
output: html_document
---

## Use JuliaCall as Julia Engine in R Markdown

To use `JuliaCall` package for julia engine in R Markdown document, just set the language of the code chunk to be `julia`.

```{julia}
## This is a julia language chunk.
## In julia, the command without ending semicolon will trigger the display
## so is JuliaCall package.
## The julia display will follow immediately after the corresponding command
## just as the R code in R Markdown.

a = sqrt(2);
a = sqrt(2)
```

### Support for `Plots.jl`

`Plots.jl` is an easy to use and powerful julia package for plotting, see <https://github.com/JuliaPlots/Plots.jl> for more detail. Before the first time using `Plots.jl`, you first need to install the package like:

```{julia eval=FALSE, include=FALSE}
import Pkg; Pkg.add("Plots")
```

After installation of the package, you can use it like this:

```{julia}
using Plots
gr()
```

```{julia}
plot(Plots.fakedata(50,5),w=3)
```

## Gurobi

Example from [link](https://gitlab.kuleuven.be/UCM/esim-code-examples/-/blob/master/MCP/MCP.ipynb)

```{julia}
using JuMP, Gurobi
CAPEX = [8000, 7000];
OPEX = [10, 100];
d = [11000, 10000, 6000, 6000, 6000, 6000, 6000, 6000, 6000, 6000, 6000, 6000, 6000, 5000, 5000];
NT = length(d);      # number timesteps
NI = length(CAPEX);  # number technologies
```

```{julia}
m = Model(Gurobi.Optimizer)

# Variables
@variable(m,g[i in 1:NI, t in 1:NT]    >= 0) # generation decision
@variable(m,y[i in 1:NI]               >= 0) # investment decision

# Constraints
@constraint(m, mu_g[i in 1:NI, t in 1:NT], -g[i,t]                          <= 0.0)
@constraint(m, mu_y[i in 1:NI],            -y[i]                            <= 0.0)
@constraint(m, mu[i in 1:NI, t in 1:NT],    g[i,t] - y[i]                     <= 0.0)
@constraint(m, dem[t in 1:NT],              
    sum(g[i,t] for i in 1:NI) - d[t] == 0.0 # energy balance
)

# minimize OPEX and CAPEX
@objective(m,Min, sum(sum( g[i,t] * OPEX[i] for t in 1:NT) + y[i] * CAPEX[i] for i in 1:NI))

optimize!(m)

y_KM = value.(y)
g_KM = value.(g)
p_KM = dual.(dem)

println("Done.")
```