### 2022-06-30 docs/Presentations/UVigo/home.Rmd
template presentación UVigo

---
title: "Cap-and-trade in Energy Capacity Investment when the planner is inattentive"
author: "Cea-Echenique, Feijoo and Muñoz"
date: "July 2022 version `r Sys.Date()`"
output: 
  beamer_presentation: 
    toc: yes
    slide_level: 2
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Motivation
## Motivation

- Bullet 1
- Bullet 2
- Bullet 3

# Model
## Model

```{r cars, echo = TRUE}
summary(cars)
```

# Results
## Results

```{r pressure}
plot(pressure)
```