### 2022-07-14 docs/Presentations/UVigo/home.Rmd
template UVigo

---
title: "Cap-and-trade in Energy Capacity Investment when the planner is inattentive"
# subtitle: "Subtitle"
author: "Cea-Echenique, Feijoo and Muñoz"
institute: "U. Andes, PUCV"
date: "July 2022<br/><br/> last update `r Sys.Date()`"
# last update: `r Sys.Date()`
# format(Sys.time(), '%e %B %Y')
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [beamer.css, beamer-fonts.css]
    nature:
      highlightStyle: github
      highlightLines: TRUE
      countIncrementalSlides: FALSE
      ratio: '16:9'
# Note: You can save the slides in .pdf format by uncommenting the line below
#knit: pagedown::chrome_print
# Note: Remark is designed such that if using incremental slides, they need to be separated by --, followed by a line break. Official documentation: https://github.com/gnab/remark/wiki/Markdown.
# Note:
## Xaringan can only include .png figures. You can include figures with:
## ```{r out.width = "60%"} # Change figure size
## knitr::include_graphics("Figures/Figure.png")
## ```
## Further instructions:
## - https://community.rstudio.com/t/size-of-image-in-xaringan-slides/21319
## - https://www.zevross.com/blog/2017/06/19/tips-and-tricks-for-working-with-images-and-figures-in-r-markdown-documents/#arguments-out.width-and-out.height-apply-to-both-existing-images-and-r-generated-figures
---

```{r setup, include=FALSE}
# Suppress the version number in the subdirectory name
options(htmltools.dir.version = FALSE)

# Define how figures are displayed
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Add option of showing tile overview by pressing "o"
xaringanExtra::use_tile_view()

# Add option to write over slides
xaringanExtra::use_scribble()

# Biblio setup
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "authoryear",
           style = "markdown",
           hyperlink = TRUE,
           dashed = FALSE
           )
# mybib <- ReadBib("~/Dropbox/Investigacion/NGOEfficiency/biblio.bib", check = FALSE)
zoterokey<- Sys.getenv("zoterokey")
zoteroID<- Sys.getenv("zoteroID")
# mybib <- ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Samuelson")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Cornes")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Austen"))  + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Gradstein")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Nocetti"))  + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Nitzan"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Cumming"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Bergstrom"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Eichberger"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Keenan"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Tamai"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Gravel"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Chetty"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Eichberger"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Chetty"))
# WriteBib(mybib, file="referencesPres.bib")
mybib <- ReadBib("referencesPres.bib")
```

# Outline

.pull-left[
- [Motivation](#motivation)
- [Why revisiting](#why)
- [Literature Review](#litrev)
- [Motivating example](#example)
- [Model](#model)
  - [Equilibrium](#equilibrium)
  <!-- - [Assumptions](#assumptions) -->
  - [Results](#results)
- [Discussion](#discussion)
]

.pull-right[
Extra
  - [Log case](#logcase)
  - [Beliefs](#resultsBeliefs)
]

???

This is the outline of the presentation. First, I would like to explain why we think we are revisiting this problem. Then, I will provide a motivation regarding some contingent discussion in order to prepare the theoretical literature revision.

Before the model is presented I will give you an example we worked in earlier versions of the manuscript that should carry the first idea we had.

Then, we revise little notations and the results in order to open a discussion to understand the meaning of the results.

I have some extra material regarding some comparative statict for some of you that may be interested in the role of beliefs.

Let us start with the argument behind revisiting this classical problem.  
---
name: motivation
# Motivation
--

- Boom of .beamer[ESG] investments:

.center[externalities, public goods, cleantech]
--

- (Technological) Uncertainty: risk aversion, asymmetric information

`r Citet(mybib, "cumming_cleantech_2016")`: uncertainty as a relevant driver

 <!-- .blue[emphasis] -->

--

- Example:

$$\downarrow\text{ Energy efficiency} + \uparrow\text{ Energy Use} \nRightarrow\downarrow\text{ Emissions}$$  
<!-- .red[emphasis] -->