### 2022-07-14 docs/Presentations/UVigo/home.Rmd
motiv y litrev

---
title: "Cap-and-trade in Energy Capacity Investment when the planner is inattentive"
# subtitle: "Subtitle"
author: "Cea-Echenique, Feijoo and Muñoz"
institute: "U. Andes, PUCV"
date: "July 2022<br/><br/> last update `r Sys.Date()`"
# last update: `r Sys.Date()`
# format(Sys.time(), '%e %B %Y')
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [beamer.css, beamer-fonts.css]
    nature:
      highlightStyle: github
      highlightLines: TRUE
      countIncrementalSlides: FALSE
      ratio: '16:9'
# Note: You can save the slides in .pdf format by uncommenting the line below
#knit: pagedown::chrome_print
# Note: Remark is designed such that if using incremental slides, they need to be separated by --, followed by a line break. Official documentation: https://github.com/gnab/remark/wiki/Markdown.
# Note:
## Xaringan can only include .png figures. You can include figures with:
## ```{r out.width = "60%"} # Change figure size
## knitr::include_graphics("Figures/Figure.png")
## ```
## Further instructions:
## - https://community.rstudio.com/t/size-of-image-in-xaringan-slides/21319
## - https://www.zevross.com/blog/2017/06/19/tips-and-tricks-for-working-with-images-and-figures-in-r-markdown-documents/#arguments-out.width-and-out.height-apply-to-both-existing-images-and-r-generated-figures
---

```{r setup, include=FALSE}
# Suppress the version number in the subdirectory name
options(htmltools.dir.version = FALSE)

# Define how figures are displayed
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Add option of showing tile overview by pressing "o"
xaringanExtra::use_tile_view()

# Add option to write over slides
xaringanExtra::use_scribble()

# Biblio setup
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "authoryear",
           style = "markdown",
           hyperlink = TRUE,
           dashed = FALSE
           )
# mybib <- ReadBib("~/Dropbox/Investigacion/NGOEfficiency/biblio.bib", check = FALSE)
zoterokey<- Sys.getenv("zoterokey")
zoteroID<- Sys.getenv("zoteroID")
# mybib <- ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Amigo")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Dewan")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="smeers"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="cea-echenique"))
# + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Gradstein")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Nocetti"))  + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Nitzan"))

# WriteBib(mybib, file="referencesPres.bib")
mybib <- ReadBib("referencesPres.bib")
```

# Outline

.pull-left[
- [Motivation](#motivation)
- [Why revisiting](#why)
- [Literature Review](#litrev)
- [Motivating example](#example)
- [Model](#model)
  - [Equilibrium](#equilibrium)
  <!-- - [Assumptions](#assumptions) -->
  - [Results](#results)
- [Discussion](#discussion)
]

.pull-right[
Extra
  - [Log case](#logcase)
  - [Beliefs](#resultsBeliefs)
]

---
name: motivation
# Motivation

- Decarbonization
  - Contrast Tax versus Cap-and-trade
  - Carbon pricing
- Environment resposability as an attention problem

---
name: literature
# Literature

- 2-stage stochastic capacity investment problems

`r Cite(mybib, c("ehrenmann_stochastic_2011","abada_multiplicity_2017","de_maere_d__aertrycke_risk_2017"))`
- Re-trading: `r Citet(mybib, "amigo_two_2021")`
- Investment segmentation: `r Citet(mybib, "cea-echenique_general_2018")`
- Rational Inattention: `r Citet(mybib, "dewan_estimating_2020")`
---
name: model
# Model

```{r timeline, out.width='80%'}
knitr::include_graphics("Fig1.png")
```

---
name: references
# References

.pull-left[
.small[
```{r refsI, echo=FALSE, results="asis"}
PrintBibliography(mybib, .opts = list(check.entries = TRUE), start=1, end=3)
```
]
]

.pull-right[
.small[

```{r refsII, echo=FALSE, results="asis"}
PrintBibliography(mybib, .opts = list(check.entries = TRUE), start=4, end=6)
```
]
]