### 2022-07-15 docs/Presentations/UVigo/home.Rmd
propuesta inicial

---
title: "Cap-and-trade in Energy Capacity Investment when the planner is inattentive"
# subtitle: "Subtitle"
author: "Cea-Echenique, Feijoo and Muñoz"
institute: "U. Andes, PUCV"
date: "July 2022<br/><br/> last update `r Sys.Date()`"
# last update: `r Sys.Date()`
# format(Sys.time(), '%e %B %Y')
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [beamer.css, beamer-fonts.css]
    nature:
      highlightStyle: github
      highlightLines: TRUE
      countIncrementalSlides: FALSE
      ratio: '16:9'
# Note: You can save the slides in .pdf format by uncommenting the line below
#knit: pagedown::chrome_print
# Note: Remark is designed such that if using incremental slides, they need to be separated by --, followed by a line break. Official documentation: https://github.com/gnab/remark/wiki/Markdown.
# Note:
## Xaringan can only include .png figures. You can include figures with:
## ```{r out.width = "60%"} # Change figure size
## knitr::include_graphics("Figures/Figure.png")
## ```
## Further instructions:
## - https://community.rstudio.com/t/size-of-image-in-xaringan-slides/21319
## - https://www.zevross.com/blog/2017/06/19/tips-and-tricks-for-working-with-images-and-figures-in-r-markdown-documents/#arguments-out.width-and-out.height-apply-to-both-existing-images-and-r-generated-figures
---

```{r setup, include=FALSE}
# Suppress the version number in the subdirectory name
options(htmltools.dir.version = FALSE)

# Define how figures are displayed
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Add option of showing tile overview by pressing "o"
xaringanExtra::use_tile_view()

# Add option to write over slides
xaringanExtra::use_scribble()

# Biblio setup
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "authoryear",
           style = "markdown",
           hyperlink = TRUE,
           dashed = FALSE
           )
# mybib <- ReadBib("~/Dropbox/Investigacion/NGOEfficiency/biblio.bib", check = FALSE)
zoterokey<- Sys.getenv("zoterokey")
zoteroID<- Sys.getenv("zoteroID")
# mybib <- ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Amigo")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Dewan")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="smeers"))
# mybib <- mybib + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="cea-echenique"))
# + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Gradstein")) + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Nocetti"))  + ReadZotero(user = zoteroID, .params=list(key = zoterokey, q="Nitzan"))

# WriteBib(mybib, file="referencesPres.bib")
mybib <- ReadBib("referencesPres.bib")
```

# Outline

.pull-left[
- [Motivation](#motivation)
- [Why revisiting](#why)
- [Literature Review](#litrev)
- [Motivating example](#example)
- [Model](#model)
  - [Equilibrium](#equilibrium)
  <!-- - [Assumptions](#assumptions) -->
  - [Results](#results)
- [Discussion](#discussion)
]

.pull-right[
Extra
  - [Log case](#logcase)
  - [Beliefs](#resultsBeliefs)
]

---
name: motivation
# Motivation

- Decarbonization
  - Contrast Tax versus Cap-and-trade
  - Carbon pricing
- Environment resposability as an attention problem

---
name: literature
# Literature

- 2-stage stochastic capacity investment problems

`r Cite(mybib, c("ehrenmann_stochastic_2011","abada_multiplicity_2017","de_maere_d__aertrycke_risk_2017"))`
- Re-trading: `r Citet(mybib, "amigo_two_2021")`
- Investment segmentation: `r Citet(mybib, "cea-echenique_general_2018")`
- Rational Inattention: `r Citet(mybib, "dewan_estimating_2020")`
---
name: model
# Model

```{r timeline, out.width='80%'}
knitr::include_graphics("Fig1.png")
```
.pull-left[
- $\bar t$ periods indexed by $T$
- $\Omega$ finite set of states of nature
]

.pull-right[
- $P$ allowance long position
- $V$ allowance short position
]
---
name: demand
# Demand

In the first stage period $t=0$,

- Exogenous demand $D(0)\in\mathbb{R}_+$.

In the second stage $t\in T:=\{1,...\bar{t}\}$, $\omega\in\Omega$
- Probability $Pr(\omega)\in[0,1]$. 
- Exogenous stochastic demand level $D(t,\omega)\in\mathbb{R}_+^{T\times\Omega}$,  $D=\left(D(0),(D(t,\omega))_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$.
---
name: producers
# Producers: first stage

$t=0$, each producer $i$ chooses 
- $Q_i(t)\in\mathbb{R}_+$ generation quantity such that their revenue is maximized price $\pi^d(0)\in\mathbb{R}_+$
- $\bar{Q}_i$ installed capacity for each producer at $t=0$
- $A_i\in\mathbb{R}_+$ carbon allowances  at a price $\pi^{a}\in\mathbb{R}_+$
- $x_i(t)\in\mathbb{R}_+$ additional capacity  with a capital expenditure of $I_i\in\mathbb{R}_+$ (The additional capacity becomes available after a predefined building time (in years))
---
name: producers2
# Producers: second stage

In the second stage, for each pair $(t,\omega)$, producer $i$ maximizes its profit by choosing 
- $Q_i(t,\omega)\in\mathbb{R}_+$ the generation level at a price $\pi^d(t,\omega)\in\mathbb{R}_+$
- $TC_i(t) \in\mathbb{R}_+$ technology change that adjusts the marginal cost of different technologies. 
- $x_i(t,\omega)\in\mathbb{R}_+$ additional capacity  with an investment cost of $TCR_i(t)\cdot I_i$, where $TCR_i(t)\in\mathbb{R}_+$ is the change in the investment cost or capital expenditure over time
- $P_i(\omega)\in\mathbb{R}_+$ long position in permits from other producers if they need to surpass the initial allowances allocation $A_i$
- $V_i(\omega)\in\mathbb{R}_+$ short position in allowances
- $\pi^v(\omega)\in\mathbb{R}_+$ permits retrade price
---
name: producers3
# Producers

The choice set of the producer $i$ over a time horizon of $\bar{t}$ years is given by: 
- an amount of capacity expansion $x_i:=\left(x_i(0),(x_i(t,\omega))_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$, 
- a production plan $Q_i:=\left(Q_{i}(0),(Q_{i}(t,\omega)_{(t,\omega)\in T\times\Omega})\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$, 
- allowances $A_i\in\mathbb{R}_+$ bought in the first period $t=0$, 
- allowances $P_i(\omega)\in\mathbb{R}_+^{\Omega}$ bought in $t=1$ for time interval $t\in[1,\bar{t}]$
- allowances $V_i(\omega)\in\mathbb{R}_+^{\Omega}$ sold in $t=1$ for time interval $t\in[1,\bar{t}]$.

Given electricity prices $\pi^d:=\left(\pi^d(0),\left(\pi^d(t,\omega)\right)_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$ and parameters $(a_i,b_i)_{i\in\{1,...N\}}\in(\mathbb{R}^2_+)^N$, $f_i(p,q)=\Big(a_i\cdot q+\frac{b_i}{2}\cdot q^{2}\Big)-p\cdot q$ revenue function.
---
name: producers4
# Producers: Objective

Defining $T_0:=\{0\}\cup T$, the optimization problem of producer $i$ is given by choosing
$(x_i,Q_i, A_i,P_i,V_i)\in\mathbb{X}:=\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right) \times\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right) \times \mathbb{R}_+\times\mathbb{R}_+^{\Omega}\times\mathbb{R}_+^{\Omega}$ given prices $(\pi^d$, $\pi^a$ , $\pi^v)\in\Pi:=\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right)\times\mathbb{R}_+\times\mathbb{R}^{\Omega}_+$ and parameters $\left((a_i,b_i),I_i, TC_i(t,\omega), TCR_i(t,\omega), CF_i,\bar{Q}_i, RP_i , \varepsilon_i\right)\in \Xi:=\mathbb{R}_+^2\times\mathbb{R}_+^7$, $\tau\in \mathbb{R}_+$ and probability $(Pr(\omega))_{\omega\in\Omega}\in\Delta:=\left\{\left(Pr(\omega)\right)_{\omega\in\Omega}\in[0,1]^K:\sum_{\omega\in\Omega}Pr(\omega)=1\right\}$ as a solution of

$$\min_{(x_i,Q_i,A_i,P_i,V_i)\in \mathbb{X}}   f_i \big( \pi^d(0),Q_i(0)\big)+ A_i \pi^{a} + I_i x_i(0)$$

$$+ \sum_{\omega} Pr(\omega)   \Bigg[ \sum_{t>0} \frac{1}{(1+R)^t} \Big[ TC_i(t)\cdot f_i \big( \pi^d(t,\omega),Q_i(t,\omega) \big)$$

$$+ TCR_i(t) \cdot I_i\cdot x_i(t,\omega) \Big] + \pi^v(\omega)\cdot \big(P_i(\omega)-V_i(\omega)\big) \Bigg] $$
---
name:prodrest
# Producers: Restrictions

$$\Big(CF_i \cdot\tau\Big)  \Bigg[\bar{Q}_i + x_i(0)+\sum_{t^{\prime}<t-lag_i} x_i(t^\prime,\omega) \Bigg] - Q_i(t,\omega) \geq 0   \quad \forall \quad i,\omega, t  > 0  \quad (\alpha_{i,\omega,t})$$

$$(CF_i\cdot\tau)\bar{Q}_i-Q_i(0)\geq 0 \qquad \forall \quad i\quad (\kappa_i)$$
$$RP_i - \bar{Q}_i - x_i(0) - \sum_{t > 0} x_i(t,\omega) \geq 0  \qquad \forall \quad i,\omega       \quad (\psi_{i,\omega})$$

$$A_{i}-V_i(\omega)\geq  0 \qquad \forall \quad i,\omega \quad (\beta_{i,\omega})$$

$$A_{i} + (P_i(\omega) - V_i(\omega))-\sum_{t>0}Q_i(t, \omega)\cdot \varepsilon_{i}-Q_i(0)\varepsilon_{i}           \geq  0  \qquad \forall \quad i, \omega     \quad (\gamma_{i,\omega})$$
---
name: auctioneer
# Auctioneer: Amigo et alii (2021)

- $\theta\in \mathbb{R}_+$ available allowances at a price $\pi^a$ in stage $t=0$. 
- The price is set such that it clears the $CO_2$ permit market among generators and the auctioneer. 
- $CAP\in\mathbb{R}_+$ carbon budget (maximum level of $CO_2$ emissions allowed in the second stage $t\in T$) and is drawn from a normal distribution, i.e., $CAP \thicksim N(\mu, \sigma^2)$.

$$Pr(\theta \geq CAP) \leq \epsilon$$

- Objective
$$\min_{\theta} -\theta \pi^{a}  + \mathcal{F}(\theta)\quad\text{s.t }\quad\phi^{-1}(\epsilon) \sigma + \mu - \theta  \geq 0$$
---
name: equilibrium
# Equilibrium

The data that parameterize the capacity investment model is given by the tuple $$\left(\tau,(Pr(\omega))_{\omega\in\Omega},\left(CAP,\mu,\sigma,\epsilon\right),\left((a_i,b_i),I_i, TC_i(t,\omega), TCR_i(t,\omega), CF_i,\bar{Q}_i, RP_i , \varepsilon_i\right)_{i\in N}\right)$$
$$\in \mathbb{R}_+\times \Delta\times\mathbb{R}_+^4\times\Xi^N,$$

**DEFINITION.** An equilibrium in the capacity investment model is a vector of prices and production decisions $$\left((\pi^{d*},\pi^{a*},\pi^{v*}),(x_i^*,Q_i^*,A_i^*,P_i^*,V_i^*)_{i\in\{1,...,N\}},\theta^*\right)\in\Pi\times\mathbb{X}^N \times \mathbb{R}_+$$ such that:
---
name: eqcond
# Equilibrium: conditions

- $(x_i^*,Q_i^*,A_i^*,P_i^*,V_i^*)$ minimizes the cost for each producer  $i\in\{1,...,N\}$
- $\theta^*$ minimizes the auctioneer cost
- Market clearing conditions are satisfied:

.pull-left.small[
- available allowances $t=0$
$$\sum_{i} A_{i}^* = \theta   \ \  (\pi^{a*})$$
- equilibrium in trading market $t>0$
$$\sum_{i} P_{i,\omega}^* = \sum_{i} V_{i,\omega}^*  \forall \ \omega  \ \ \left(\pi^{v*}(\omega)\right)$$
]

.pull-right.small[

- fulfillment of the demand --first stage
$$\sum_{i} Q_i(0)^* = D(0), \ \ (\pi^{d*}(0))$$
- fulfillment of the demand --second stage
$$\sum_{i} Q_i(t,\omega)^* = D(t,\omega),  \forall \ \omega, t \quad(\pi^{d*}(t,\omega))$$ 
]  
---
name: attention
# Rational Inattention

Following `r Citet(mybib, "dewan_estimating_2020")` we consider a quadratic cost function

$$C(P)=\begin{cases}0,&P\leq d\\c(P-d)^2,&P>d\end{cases},$$

where $P\in[0,1]$ is a performance metric, $c$ an associated marginal cost and $d$ the performance level wiht public information.

We drop the chance constraint approach, by using

- $\min_{(\theta,P)\in[0,(1+\epsilon)CAP]\times[0,1]}-\theta\pi^aP+c(P-d)^2$ profit oriented
- $\min_{\theta\in[0,CAP]}-\theta\pi^a+c\left(1-\frac{|\theta-CAP|}{CAP}-d\right)^2$ welfare oriented
---
name: results
# Results: Profit oriented

- Calibration wrt `r Citet(mybib, "amigo_two_2021")`
  - CAP: $100 MtCO_2$
  - $\pi^a=315.83$
  - $\epsilon=0.202$ Andres et alii (2014)
$$\Rightarrow c=920M$$

---
name: remarks
# Final Remarks

- Multiple solutions and comparative statics

---
name: references
# References

.pull-left[
.small[
```{r refsI, echo=FALSE, results="asis"}
PrintBibliography(mybib, .opts = list(check.entries = TRUE), start=1, end=3)
```
]
]

.pull-right[
.small[

```{r refsII, echo=FALSE, results="asis"}
PrintBibliography(mybib, .opts = list(check.entries = TRUE), start=4, end=6)
```
]
]