### 2023-07-13 docs/Presentations/IFORS/home.qmd

---
title: Inattention cost of energy
subtitle: Cap-and-trade in Energy Capacity Investment when the planner is inattentive
format:
  revealjs:
    slide-number: true
    show-slide-number: all
editor: visual
author: "Cea-Echenique, Feijoo and Muñoz"
institute: "UAndes, PUCV"
date: "July 2023<br/><br/> last update `r Sys.Date()`"
html-math-method: katex
bibliography: references.bib
---

##  {background-image="images/frontSlide.png"}

## Outline

-   [Motivation](#motivation)
-   [Model](#model)
-   [Results](#results)

## Motivation {#motivation}

-   Decarbonization
    -   Contrast Tax versus Cap-and-trade
    -   Carbon pricing
-   Environment responsability as an attention problem

## Literature

-   2-stage stochastic capacity investment problems

[@ehrenmann_stochastic_2011; @de_maere_d__aertrycke_risk_2017; @abada_multiplicity_2017]

-   Re-trading: @amigo_two_2021

-   Investment segmentation: @cea-echenique_general_2018

-   Rational Inattention: @dewan_estimating_2020

## Model {#model}

![](images/Fig1.png){fig-align="center"}

::: columns
::: {.column width="50%"}
-   $\bar t$ periods indexed by $T$
-   $\Omega$ finite set of states of nature
:::

::: {.column width="50%"}
-   $P$ allowance long position
-   $V$ allowance short position
:::
:::

## Demand {.smaller}

In the first stage period $t=0$,

-   Exogenous demand $D(0)\in\mathbb{R}_+$.

In the second stage $t\in T:=\{1,...\bar{t}\}$, $\omega\in\Omega$

-   Probability $Pr(\omega)\in[0,1]$.
-   Exogenous stochastic demand level $D(t,\omega)\in\mathbb{R}_+^{T\times\Omega}$,

$$D=\left(D(0),(D(t,\omega))_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}.$$

## Producers: first stage {.smaller}

$t=0$, each producer $i$ chooses

-   $Q_i(t)\in\mathbb{R}_+$ generation quantity such that their revenue is maximized price $\pi^d(0)\in\mathbb{R}_+$
-   $\bar{Q}_i$ installed capacity for each producer at $t=0$
-   $A_i\in\mathbb{R}_+$ carbon allowances at a price $\pi^{a}\in\mathbb{R}_+$
-   $x_i(t)\in\mathbb{R}_+$ additional capacity with a capital expenditure of $I_i\in\mathbb{R}_+$ (The additional capacity becomes available after a predefined building time (in years))

## Producers: second stage {.smaller}

In the second stage, for each pair $(t,\omega)$, producer $i$ maximizes its profit by choosing

-   $Q_i(t,\omega)\in\mathbb{R}_+$ the generation level at a price $\pi^d(t,\omega)\in\mathbb{R}_+$
-   $TC_i(t) \in\mathbb{R}_+$ technology change that adjusts the marginal cost of different technologies.
-   $x_i(t,\omega)\in\mathbb{R}_+$ additional capacity with an investment cost of $TCR_i(t)\cdot I_i$, where $TCR_i(t)\in\mathbb{R}_+$ is the change in the investment cost or capital expenditure over time
-   $P_i(\omega)\in\mathbb{R}_+$ long position in permits from other producers if they need to surpass the initial allowances allocation $A_i$
-   $V_i(\omega)\in\mathbb{R}_+$ short position in allowances
-   $\pi^v(\omega)\in\mathbb{R}_+$ permits retrade price

## Producers {.smaller}

The choice set of the producer $i$ over a time horizon of $\bar{t}$ years is given by:

-   an amount of capacity expansion $x_i:=\left(x_i(0),(x_i(t,\omega))_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$,
-   a production plan $Q_i:=\left(Q_{i}(0),(Q_{i}(t,\omega)_{(t,\omega)\in T\times\Omega})\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$,
-   allowances $A_i\in\mathbb{R}_+$ bought in the first period $t=0$,
-   allowances $P_i(\omega)\in\mathbb{R}_+^{\Omega}$ bought in $t=1$ for time interval $t\in[1,\bar{t}]$
-   allowances $V_i(\omega)\in\mathbb{R}_+^{\Omega}$ sold in $t=1$ for time interval $t\in[1,\bar{t}]$.

Given electricity prices $\pi^d:=\left(\pi^d(0),\left(\pi^d(t,\omega)\right)_{(t,\omega)\in T\times\Omega}\right)\in\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}$ and parameters $(a_i,b_i)_{i\in\{1,...N\}}\in(\mathbb{R}^2_+)^N$, $f_i(p,q)=\Big(a_i\cdot q+\frac{b_i}{2}\cdot q^{2}\Big)-p\cdot q$ revenue function.

## Producers: Objective {.smaller}

::: panel-tabset
### Variables and Parameters

Defining $T_0:=\{0\}\cup T$, the optimization problem of producer $i$ is given by choosing

-   $(x_i,Q_i, A_i,P_i,V_i)\in\mathbb{X}:=\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right) \times\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right) \times \mathbb{R}_+\times\mathbb{R}_+^{\Omega}\times\mathbb{R}_+^{\Omega}$
-   given prices $(\pi^d$, $\pi^a$ , $\pi^v)\in\Pi:=\left(\mathbb{R}_+\times\mathbb{R}_+^{T\times\Omega}\right)\times\mathbb{R}_+\times\mathbb{R}^{\Omega}_+$,
-   parameters $\left((a_i,b_i),I_i, TC_i(t,\omega), TCR_i(t,\omega), CF_i,\bar{Q}_i, RP_i , \varepsilon_i\right)\in \Xi:=\mathbb{R}_+^2\times\mathbb{R}_+^7$, $\tau\in \mathbb{R}_+$
-   and probability $(Pr(\omega))_{\omega\in\Omega}\in\Delta:=\left\{\left(Pr(\omega)\right)_{\omega\in\Omega}\in[0,1]^K:\sum_{\omega\in\Omega}Pr(\omega)=1\right\}$ as a solution of

### Objective function

$$\min_{(x_i,Q_i,A_i,P_i,V_i)\in \mathbb{X}}   f_i \big( \pi^d(0),Q_i(0)\big)+ A_i \pi^{a} + I_i x_i(0)$$

$$+ \sum_{\omega} Pr(\omega)   \Bigg[ \sum_{t>0} \frac{1}{(1+R)^t} \Big[ TC_i(t)\cdot f_i \big( \pi^d(t,\omega),Q_i(t,\omega) \big)$$

$$+ TCR_i(t) \cdot I_i\cdot x_i(t,\omega) \Big] + \pi^v(\omega)\cdot \big(P_i(\omega)-V_i(\omega)\big) \Bigg] $$

### Restrictions

$$\Big(CF_i \cdot\tau\Big)  \Bigg[\bar{Q}_i + x_i(0)+\sum_{t^{\prime}<t-lag_i} x_i(t^\prime,\omega) \Bigg] - Q_i(t,\omega) \geq 0,\forall \, i,\omega, t  > 0  \, (\alpha_{i,\omega,t})$$

$$(CF_i\cdot\tau)\bar{Q}_i-Q_i(0)\geq 0 \qquad \forall \quad i\quad (\kappa_i)$$ $$RP_i - \bar{Q}_i - x_i(0) - \sum_{t > 0} x_i(t,\omega) \geq 0  \qquad \forall \quad i,\omega       \quad (\psi_{i,\omega})$$

$$A_{i}-V_i(\omega)\geq  0 \qquad \forall \quad i,\omega \quad (\beta_{i,\omega})$$

$$A_{i} + (P_i(\omega) - V_i(\omega))-\sum_{t>0}Q_i(t, \omega)\cdot \varepsilon_{i}-Q_i(0)\varepsilon_{i}           \geq  0  \qquad \forall \quad i, \omega     \quad (\gamma_{i,\omega})$$
:::

## Auctioneer: Amigo et alii (2021) {.smaller}

-   $\theta\in \mathbb{R}_+$ available allowances at a price $\pi^a$ in stage $t=0$.
-   The price is set such that it clears the $CO_2$ permit market among generators and the auctioneer.
-   $CAP\in\mathbb{R}_+$ carbon budget (maximum level of $CO_2$ emissions allowed in the second stage $t\in T$) and is drawn from a normal distribution, i.e., $CAP \thicksim N(\mu, \sigma^2)$.

$$Pr(\theta \geq CAP) \leq \epsilon$$

-   Objective $$\min_{\theta} -\theta \pi^{a}  + \mathcal{F}(\theta)\quad\text{s.t }\quad\phi^{-1}(\epsilon) \sigma + \mu - \theta  \geq 0$$

## Equilibrium {.smaller}

::: panel-tabset
### Definition

The data that parameterize the capacity investment model is given by the tuple $$(\tau,(Pr(\omega))_{\omega\in\Omega},\left(CAP,\mu,\sigma,\epsilon\right),((a_i,b_i),I_i, TC_i(t,\omega), TCR_i(t,\omega), CF_i,$$

$$\bar{Q}_i, RP_i , \varepsilon_i)_{i\in N})\in \mathbb{R}_+\times \Delta\times\mathbb{R}_+^4\times\Xi^N,$$

**DEFINITION.** An equilibrium in the capacity investment model is a vector of prices and production decisions $$\left((\pi^{d*},\pi^{a*},\pi^{v*}),(x_i^*,Q_i^*,A_i^*,P_i^*,V_i^*)_{i\in\{1,...,N\}},\theta^*\right)\in\Pi\times\mathbb{X}^N \times \mathbb{R}_+$$ such that:

### Conditions

$(x_i^*,Q_i^*,A_i^*,P_i^*,V_i^*)$ minimizes the cost for each producer $i\in\{1,...,N\}$ - $\theta^*$ minimizes the auctioneer cost - Market clearing conditions are satisfied:

::: columns
::: {.column width="50%"}
-   available allowances $t=0$ $$\sum_{i} A_{i}^* = \theta   \ \  {\color{blue}{(\pi^{a*})}}$$
-   equilibrium in trading market $t>0$ $$\sum_{i} P_{i,\omega}^* = \sum_{i} V_{i,\omega}^*  \forall \ \omega  \ \ {\color{blue}{\left(\pi^{v*}(\omega)\right)}}$$
:::

::: {.column width="50%"}
-   fulfillment of the demand: first stage $$\sum_{i} Q_i(0)^* = D(0), \ \ {\color{blue}{(\pi^{d*}(0))}}$$
-   fulfillment of the demand: second stage $$\begin{array}{ll}\sum_{i} Q_i(t,\omega)^* = D(t,\omega),  &\forall \ \omega, t\\ & {\color{blue}{(\pi^{d*}(t,\omega))}}\end{array}$$
:::
:::
:::

## Rational Inattention {.smaller}

Following @dewan_estimating_2020 we consider a quadratic cost function

$$C(P)=\begin{cases}0,&Perf\leq d\\c(Perf-d)^2,&Perf>d\end{cases},$$

where $Perf\in[0,1]$ is a performance metric, $c$ an associated marginal cost and $d$ the performance level wiht public information.

We drop the chance constraint approach, by using

-   $\min_{(\theta,Perf)\in[0,(1+\epsilon)CAP]\times[0,1]}-\theta\pi^aPerf+c(Perf-d)^2$ profit oriented
-   $\min_{\theta\in[0,CAP]}-\theta\pi^a+c\left(1-\frac{|\theta-CAP|}{CAP}-d\right)^2$ welfare oriented

## Results: Profit oriented {#results .smaller}

::: columns
::: {.column width="50%"}
-   Calibration wrt @amigo_two_2021
    -   CAP: $100 MtCO_2$
    -   $\pi^a=315.83$
    -   $\epsilon=0.202$ Andres et alii (2014) $$\Rightarrow c=920M$$
:::

::: {.column width="50%"}
$CAP=100$

| Perf  | $\theta$    | $\frac{\theta}{CAP}$ | $\pi^a$ |
|-------|-------------|----------------------|---------|
| 0.798 | 120,200,000 | 1.20                 | 270.13  |
| 0.8   | 120,200,000 | 1.20                 | 270.13  |
| 0.85  | 113,654,321 | 1.14                 | 25.75   |
| 0.9   | 113,568,965 | 1.14                 | 93.65   |
| 0.95  | 118,470,479 | 1.18                 | 188.86  |
| 0.99  | 114,916,396 | 1.15                 | 298.11  |
| 1     | 118,332,486 | 1.18                 | 317.24  |
:::
:::

## Results: multiple CAPs, $Perf=0.8$ {.smaller}

| CAP           | $\theta$    | $\frac{\theta}{CAP}$ | $\pi^a$ |
|---------------|-------------|----------------------|---------|
| 100,000,000   | 120,200,000 | 1.20                 | 270.13  |
| 200,000,000   | 240,400,000 | 1.20                 | 60.57   |
| 300,000,000   | 360,600,000 | 1.20                 | 119.47  |
| 400,000,000   | 454,798,037 | 1.14                 | 95.83   |
| 500,000,000   | 601,000,000 | 1.20                 | 79.31   |
| 600,000,000   | 605,279,899 | 1.01                 | 66.57   |
| 700,000,000   | 605,279,899 | 0.86                 | 56.37   |
| 800,000,000   | 961,600,000 | 1.20                 | 152.50  |
| 900,000,000   | 605,279,899 | 0.67                 | 42.26   |
| 1,000,000,000 | 605,279,899 | 0.61                 | 35.98   |

## Results: Profit oriented, allowances distributions

::: columns
::: {.column width="50%"}
![](images/diststage1.png)
:::

::: {.column width="50%"}
![](images/diststage2.png)
:::
:::

## Results: Welfare oriented {.smaller}

::: columns
::: {.column width="50%"}
-   Calibration wrt @amigo_two_2021
    -   CAP: $100 MtCO_2$
    -   $\pi^a=315.83$

$$\Rightarrow c=9920M$$
:::

::: {.column width="50%"}
| CAP (M) | $\theta$    | Perf  | $\frac{\theta}{CAP}$ | $\pi^a$ |
|---------|-------------|-------|----------------------|---------|
| 100     | 144,944,410 | 0.798 | 1.449                | 232.82  |
| 200     | 289,888,820 | 0.798 | 1.449                | 141.56  |
| 300     | 434,833,230 | 0.798 | 1.449                | 103.33  |
| 400     | 579,777,640 | 0.798 | 1.449                |         |
| 500     | 653,754,577 | 0.905 | 1.308                | 37.27   |
| 600     | 619,939,821 | 0.999 | 1.033                | 40.43   |
| 700     | 661,258,561 | 0.997 | 0.945                | 36.42   |
| 800     | 610,913,181 | 0.944 | 0.764                | 31.28   |
| 900     | 609,127,279 | 0.896 | 0.677                | 31.21   |
| 1,000   | 698,676,376 | 0.909 | 0.699                | 30.00   |
:::
:::

## Final Remarks

-   Respect of aggregate demand law
-   Unclear relation of inattention
-   Multiple solutions and comparative statics

##  {background-image="images/endSlide.png"}

## References