### 2023-08-18 docs/Apuntes/cnt_v2.tex

\documentclass[11pt, letterpaper]{article}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[gen]{eurosym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{subcaption} 
\usepackage[title]{appendix}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
%\usepackage{hyperref}

\captionsetup{font=footnotesize}

\newcommand\boldred[1]{\textcolor{red}{\textbf{#1}}}
\newcommand{\seba}{\textcolor[rgb]{0,1,1}}

\title{Cap-and-trade Model for the Chilean Energy Market}
\author{P\'ia Amigo, Sebasti\'an Cea, Felipe Feijoo}
\date{\today}

\begin{document}
\maketitle

\section{Simple Model}\label{model}

Let us consider a market of $i$ producers, each of them minimizing the cost of operation and investment. We will assume perfect competition among the producers. Our model will assume two basic stages: operation at present (with investment for the future) and the future, at the spot market and trading of emission allowances.

\smallskip

At first stage, we consider the current operating plant and the cost of the operation $C_i$, and the quantity produce $Q_i^{\tau}$ with $\tau=1$. The price of sale (that satisfies the demand) at this stage is $\pi^{d,\tau}_i$ with $\tau=1$. The maximum quantity of production at current stage (initial) per producer is $\bar{Q}_i$. Each producer buys a certain amount of carbon $A_i$ at a price $\pi^{a}$ from an auctioner of regulatory agent.  The auctioneer will try to maximize the revenue obtained from the sold allowances. At this stage, the producer $i$ decides the amount $x_i$ of capacity expansion at a cost $I_i$ for the uncertain future.
\smallskip

At second stage, in the uncertain future, the producer wants to operate in the most efficient way at the spot market. In addition,  we will consider the existence of a permit trading market, where producers can purchase permits from other producers if they need to surpass the emission allowances, or sell the unused permits if their emissions are below allowances. The producer will evaluate the cost of the uncertain future as risk neutral agents, i.e., as an expectation value and we will assume that the probability is a physical probability, which may be obtained from real world data.
\smallskip

The maximum budget for emissions is the  parameter CAP, which in turn, follows a normal distribution, i.e., $CAP \thicksim N(\mu, \sigma^2)$. This CAP is determined by the auctioneer and  accounts for the total emissions throughout the whole period considered, that is, does not depend of $\tau$.

\smallskip
The amount of total allowances provided by the auctioneer $\theta$ is an endogenous random variable in our model, and it must satisfied the condition $\sum_i A_i = \theta$. 
Since $\theta$ must not surpass the budget for emissions, we consider a low margin (or risk) $R$ for this to happen, i.e.,
\begin{equation}
    P (\theta \geq CAP) \leq R
\end{equation}

and since the variable CAP follows a normal distribution with mean $\mu$ and variance $\sigma^2$, we can write the equation above as
\begin{equation}
    \theta \leq \phi^{-1}(R) \sigma^2 + \mu
\end{equation}

where $\phi^{-1}$ is the inverse of the cumulative function of the normal distribution.

\hspace{0.5cm}

\begin{flushleft}
\textbf{Sets}
\begin{itemize}
    \item[] $i$: number of producers
    \item[] $\omega$: possible scenarios
    \item[] $\tau$: model periods

\end{itemize}

\hspace{0.5cm}

\textbf{Parameters}
\begin{itemize}
   \item[] $\bar{Q}_i$: maximum operation capacity
   \item[] $I_i$: expansion cost
   \item[] $C_i$: operation cost
   \item[] $D$: total demand (exogenous)
   \item[] $CAP$: maximum budget of emissions 
   \item[] $R$: margin for total emission allowances
 \item[] $\varepsilon_{i}$: Emissions factor of producer $i$
\end{itemize}
\hspace{0.5cm}

\textbf{Variables}
\begin{itemize}
    \item[] $Q^{\tau}_i$: produced quantity
    \item[] $A_i$: emission allowances
    \item[] $P^{\tau}_i$: purchased permits
    \item[] $V^{\tau}_i$: sold permits
    \item[] $\pi^{a}_i$: dual to CAP constraint
    \item[] $\pi^{v}_i$: dual to trading equilibrium
    \item[] $\pi^{d}_i$: dual to fulfilment of the demand
    \item[] $x_i$: capacity expansion decision
    \item[] $\theta$: emission permits available in the market
    \end{itemize}

\hspace{0.5cm}

\textbf{Equations}
\smallskip
The model below is defined for each $i \in \{ 1,...,n\}$. Each producer minimizes cost as follows

\begin{align} \label{eq:simplemodel}
    \min_{x,Q,A,P,V} & C_iQ_i^{1} + A_i \pi^{a} + I_i x_i \nonumber\\
    & + \sum_{\Omega} prob_\omega \Big[ \sum_{\tau>1} (C_i-\pi_{\omega}) Q_{i,\omega}^{\tau}- V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i} P_{i,j,\omega}\pi_{j,w}^v\Big]\\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \bar{Q_i} + x_{i} - Q_{i,\omega}^{\tau} & \geq 0  & \forall & \quad \omega, \tau  > 1 & \quad (\alpha_{i,\omega,\tau}) \\
    \bar{Q_i}-Q_{i}^{\tau} & \geq 0  &  & \quad \tau  =1 & \quad (\kappa_i) \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall & \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall & \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_{i, \omega}^{\tau} & \geq  0   & \forall & \quad \omega, \tau & \quad (\delta_{i,\omega})
\end{align}

\smallskip
The auctioneer maximizes its revenue from the allowances market; we write the problem as a minimization program as follows
\begin{align}
    \min_{\theta} & -\theta \pi^{a} \\
    \textrm{s.t \ } & \theta \leq \phi^{-1}(R) \sigma^2 + \mu \\
    &\theta \geq 0
\end{align}

\smallskip

Finally, we have three market clearing constraints equations for this market

\begin{align}
\textrm{(available allowances)}: &  \ \   \sum_{\tau} \sum_{i} A^{\tau}_{i,\omega} = \theta   & \forall \ \omega & \ \  (\pi^{a})\\
\textrm{(equilibrium in trading market)}: &   \ \  \sum_{i} P_{i,\omega} = \sum_{i} V_{i,\omega} & \forall \ \omega & \ \ (\pi^{v}) \\
\textrm{(fulfillment of the demand)}:  &   \ \  \sum_{i} Q^{\tau}_{i,\omega} = D^{\tau}_{\omega}, & \forall \ \omega, \tau & \ \ (\pi^{d})
\end{align}

\end{flushleft}

\subsection{KKT conditions}

The Lagrangian function for the producer problem for each $i \in \{ 1,...,n\}$ is defined as follows:

\begin{multline}
    \mathcal{L}_i(x,Q,A,P,V) = C_i Q_i^{1} + A_i\pi^{a} + I_i x_i + \sum_{\Omega} prob_\omega \Big[ \sum_{\tau>1} (C_i-\pi_{\omega}) Q_{i,\omega}^{\tau} - V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i} \nonumber P_{i,j,\omega}\pi_{j,w}^v\Big]\\ + \kappa_{i}\Big[Q_{i}^{1} - \bar{Q_i}\Big] +\sum_{\omega,\tau>1} \alpha_{i,\omega,\tau}\Big[Q_{i,\omega}^{\tau} - (\bar{Q_i}+ x_i)\Big] + \sum_{\omega}\beta_{i,\omega}\Big[V_{i,\omega}-A_i\Big] \\
    + \sum_{\omega}\gamma_{i,\omega} \Big[-A_{i} - \sum_{j \neq i}P_{i,j,\omega} + V_{i,\omega}+\sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i}+Q_i^{1}\varepsilon_{i}\Big] - \sum_{\omega, \tau>1}\delta_{i,\omega,\tau} Q_{i,\omega}^{\tau} - \lambda_{i}\Big[Q^{1}_{i}\Big]
\end{multline}

where $\alpha_{i,\omega,\tau}$, $\kappa_i$, $\beta_{i,\omega}$, $\gamma_{i,\omega}$ and $\delta_{i,\omega}$ and $\lambda_i$ are the lagrange multipliers of the constraints.

\smallskip

With the Lagrangian set up, we can now find de KKT conditions for each of the producers $i \in \{ 1,...,n\}$ by making $\nabla \mathcal{L}_{i}=0$, we obtain:

\begin{align}
    I_i - \sum_{\omega,\tau>1} \alpha_{i,\omega,\tau} & = 0 & \forall i  &  \qquad (x_i)\\
    C_{i} + \kappa_i - \lambda_{i} + \sum_{\omega} \gamma_{i,\omega} & = 0 & \forall i  &  \qquad (Q^{1}_{i})\\
    prob_\omega (C_i-\pi_{\omega}) + \alpha_{i,\omega,\tau} + \gamma_{i,\omega} \varepsilon_{i}-\delta_{i,\omega,\tau} & =0 & \forall i, \omega, \tau>1 &  \qquad (Q_{i,\omega}^{\tau})\\
    \lambda - \sum_{\omega}\beta_{i,\omega} - \sum_{\omega}\gamma_{i,\omega} & =0 & \forall i & \qquad (A_{i}) \\
    -prob_\omega \pi^v_{i,\omega} + \beta_{i,\omega}  + \gamma_{i,\omega} & =0 & \forall i, \omega & \qquad (V_{i,\omega}) \\
    prob_\omega \pi^v_{j,\omega} -\gamma_{i,\omega} & = 0 & \forall i, \omega & \qquad (P_{i,j,\omega})
\end{align}

\smallskip

\begin{flushleft}
Primal feasibility
\end{flushleft}

\begin{align}
    -Q_{i,\omega}^{\tau} + \bar{Q_i} + x_i & \geq 0 & \qquad (\alpha_{i,\omega,\tau}) \\
    -\bar{Q_i} + Q^{\tau}_i & \geq 0 & \qquad (\kappa_i)\\
     A_{i} - V_{i,\omega} & \geq 0 & \qquad (\beta_{i,\omega}) \\
     A_{i} + P_{i,\omega} - V_{i,\omega} - \sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i} & \geq 0 & \qquad (\gamma_{i,\omega}) \\
    Q_{i,\omega}^{\tau} & \geq 0 & \qquad (\delta_{i,\omega})
\end{align}

\smallskip

\begin{flushleft}
Complementary slackness\\

\textbf{FALTAN los For all i,w,t...etc}
\end{flushleft}

\begin{align}
    (-Q_{i,\omega}^{\tau} + \bar{Q_i} + x_i)\alpha_{i,\omega,\tau} & = 0 \\
    ( -\bar{Q_i} + Q^{\tau}_i)\kappa_i & = 0 \\
    (V_{i,\omega} - A_{i})\beta_{i,\omega} & = 0 \\
    ( A_{i} + P_{i,\omega} - V_{i,\omega} - \sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i})\gamma_{i,\omega} & = 0 \\
    Q_{i,\omega}^{\tau}\delta_{i,\omega} & = 0
\end{align}

\smallskip

\begin{flushleft}
Dual Feasibility
\end{flushleft}

\begin{align}
    \alpha_{i,\omega,\tau} & \geq 0 \\
    \kappa_i & \geq 0 \\
    \beta_{i,\omega} & \geq 0 \\
    \gamma_{i,\omega} & \geq 0 \\
    \delta_{i,\omega} & \geq 0 
\end{align}

For the Auctioneer, the Lagrangian function is

\begin{equation}
    \mathcal{L}(\theta)= -\theta \pi^{a} - \eta (\phi^{-1}(R) \sigma^2 + \mu - \theta) - mu\theta
\end{equation}

where $\eta$ is the lagrange multiplier. Thus, we can obtain the KKT conditions for the auctioneer problem.

\begin{align}
    -\pi^{a} +\eta -\mu & = 0 \\
    \theta & \geq  0 \\
    (\phi^{-1}(R) \sigma^2 + \mu - \theta) \eta & =  0\\
    \eta & \geq  0
\end{align}

This model has been implemented in GAMS. We have considered a two stage problem, where first stage consists in the year range $2017-2018$ and second stage consists in the range $2019-2037$. We have considered each type of energy generation as a producer. The types of technology considered, according to data from the Comisi\'on Nacional de Energ\'ia are: Biogas, Biomasa, Carbon, Cogenercion, Eolica, FuelOil, Gas, Geotermia, Hydro, PetroleoDiesel, Solar. Also, data for Capacity, Cost of investment, demand have been obtained from the same database.

\section{A slightly more complicated model}

We make things a little more general by considering more detailed aspects:

\begin{itemize}
    \item We consider that at first stage, the plant is operating and fulfiling the demand at first stage, at a price $\pi^{d,1}$
    \item Producers can invest in capacity expansion not only at first stage, but also at any point during second stage. 
    \item Producers can change their plant's technology to maintain the production
\end{itemize}

We modify Equation \ref{eq:simplemodel} taken into account the considerations described above; thus, the problem takes the following form:

\begin{align}
    \min_{x,Q,A,P,V} & (C_i-\pi^{1}) Q_i^{1} + A_i \pi^{a} + I_i x_i^{1} + \sum_{\Omega} prob_\omega \Big[ \sum_{\tau>1} (TC_{i,\tau}\cdot C_i-\pi_{\omega}) Q_{i,\omega}^{\tau}- \nonumber \\ 
    &V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i}  P_{i,j,\omega}\pi_{j,w}^v+ \sum_{\tau > 1} I_ix_{i,\tau,w}\Big]\\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \bar{Q_i} + \sum_{\tau^{t}\leq\tau} x_{i,\tau^{t},w} - Q_{i,\omega}^{\tau} & \geq 0  & \forall  \quad \omega, \tau  > 1 & \quad (\alpha_{i,\omega,\tau})\\
    \bar{Q_i}-Q_{i}^{\tau} & \geq 0  &  \quad \tau  =1 & \quad (\kappa_i) \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall  \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{\tau>1}Q_{i, \omega}^{\tau}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_{i, \omega}^{\tau} & \geq  0   & \forall  \quad \omega, \tau & \quad (\delta_{i,\omega})
\end{align}

\section{Model -- Version 6}

\begin{center}
\begin{tabular}{ l l  } 
 \hline
 \textbf{Set} &  \textbf{Explanation} \\ 
 \hline
 \hline
 $i$ & Producers (Technologies) \\  
 $\omega$ & Possible scenarios \\ 
 $t$ & Periods \\
 \hline
\end{tabular}
\end{center}

For this model, each producer corresponds to a different technology present in the chilean electricity generation system, i.e., Biomasa, Carbon, Eolica, Gas, Geotermica, HidroEmbalse, HidroPasada, HidroMiniPasada, PetroleoDiesel, Solar.

\begin{center}
\begin{tabular}{  l l l l  } 
 \hline
 \textbf{Parameter} & \boldred{Suggested} & \textbf{Units} & \textbf{Explanation} \\ 
 \hline
 \hline
 $I_i$ & & USD/MW & Expansion cost per technology $i$ \\  
 $\bar{Q}_i$ & \boldred{$\bar{Y}_i$} & MW & Maximum operation capacity per technology $i$  \\ 
 $C_i$ &  & USD/MWh & Operation cost per technology $i$ \\
 $D^{1}$ & & MWh & Total demand in first stage \\
 $D^{t}_{\omega}$ & & MWh & Total demand in period $t > 1$ \\
 $\varepsilon_i$ & & tCO$_{2}$/MWh  & Emission factor of technology $i$\\
 Prob$_\omega$ & \boldred{$\Pi_\omega$} & & Physical probability of scenario $\omega$ \\
 $\mu$ & & tCO$_{2}$ & Mean of the normal distribution of the CAP \\
  $\sigma$ & & tCO$_{2}$ & Standard deviation of the normal distribution of the CAP \\
  $\epsilon$ & &  & Margin for total emission allowances \\
  $R$ & \boldred{df} & & Discount rate \\
  $TC_i$ & & & Change in the cost of operation per technology \\
  $TCR_i$ & & & Change in the investment cost per technology \\
  $CF_i$ & & & Capacity factor per technology\\
  $\tau$ & & hours&  Number of hours in a year, $\tau=8760$ hours\\
  {\color{blue} $RP_i$ } & & {\color{blue} MW} & {\color{blue} Resource potential per technology }\\
\hline
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{  l l l l  } 
 \hline
 \textbf{Variable} & \boldred{Suggested} & \textbf{Units} & \textbf{Explanation} \\ 
 \hline
 \hline
 $Q^{1}_i$ & \boldred{$Y^{1}_{i}$} & MWh & produced quantity in period $t =1  $ for producer $i$\\
 $Q^{t}_{i,\omega}$ & \boldred{$Y^{t}_{i,\omega}$} & MWh & produced quantity in period $t > 1 $ for producer $i$ in scenario $\omega$ \\  
 $A_i$ & \boldred{$\theta_i$} & tCO$_2$ & Emission allowances purchased by producer $i$ \\ 
 $P_{i,j,\omega}$ & \boldred{$B_i$}& tCO$_2$ & Purchased permits in trading market by producer i  \\
 $V_{i,\omega}$ & & tCO$_2$ & Sold permits in trading markets \\
 $\pi^{a}$ & \boldred{$P^{a}$} & USD/tCO$_2$  & Price of the allowances offered by the auctioneer (dual to CAP constraints)\\
 $\pi^{v}_{i,\omega}$ & \boldred{$P^{v}_{i,\omega}$} & USD/tCO$_2$  & Price of the permits in trading market (dual to trading equilibrium)\\
  $\pi^{d,1}$ & \boldred{$P^{d,1}$} & USD/MWh  & Price of electricity (dual to fulfilment of the demand --first stage)\\
 $\pi^{d,t}_{\omega}$ & \boldred{$P^{d,t}$} & USD/MWh  & Price of electricity (dual to fulfilment of the demand)\\
 $x^{1}_i$ & & MW  & Capacity expansion decision for period $t=1$ of producer $i$\\
 $x^{t}_{i,\omega}$ & & MW  & Capacity expansion decision for period $t$ of producer $i$ in scenario $\omega$\\
  $\theta$ & \boldred{$\Theta$} & tCO$_2$ & Emission allowances available in the market by the auctioneer\\
\hline
\end{tabular}
\end{center}

\vspace{0.5cm}

\underline{Producer's problem} (each producer is a different technology):

\begin{align}
\min_{x,Q,A,P,V} & \underbrace{{\color{red}\Big(a_iQ_{1,i}+\frac{b}{2}Q_{1,i}^{2}\Big)-\pi^{d,1} Q_{1,i}}}_{\seba{\left(C_t^{comb}\cdot C^{esp}+CVnC\right)Q_{i,t}+\underbrace{FixedCost_{i,t}}_{\frac{\alpha\cdot I\cdot x}{algo}}}} + A_i \pi^{a} + I_i x_i^{1} + \sum_{\Omega} prob_\omega   \Bigg( \sum_{t>1} \frac{1}{(1+R)^t} \Big[ (TC_{i,\tau}\cdot {\color{red}(a_i Q_{i,t,\omega}+\frac{b}{2}Q_{i,t,\omega}^{2})}\nonumber \\
    & {\color{red}-\pi^{d,2}_{\omega} Q_{i,t,\omega}} + \sum_{t > 1} TCR_i \cdot I_ix^{t}_{i,w} \Big] - V_{i,\omega}\pi_{i,w}^v+\sum_{j\neq i}  P_{i,j,\omega}\pi_{j,w}^v\Bigg)  \label{prodFO} \\
     \textrm{s.t \ } \nonumber
\end{align}
\begin{align}
    \Big(CF_i \cdot\tau\Big)  \Bigg[\bar{Q_i} + \sum_{t^{\prime}<t_{y_i}} x^{t^{\prime}}_{i,w} + x_i^{1}+ \bar{Q}_{i,t} \Bigg] - Q_{i,\omega}^{t} & \geq 0  & \forall  \quad \omega, t  > 1 & \quad (\alpha_{i,\omega,t}) \label{const.1}\\
    \Big(CF_i\cdot\tau \Big)\bar{Q_i}-Q_{i}^{t} & \geq 0  &  \quad t  =1 & \quad (\kappa_i) \label{const.2} \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall  \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{t>1}Q_{i, \omega}^{t}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_i^{1} & \geq  0 & \quad & \quad (\lambda_i)\\
 Q_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\delta_{i,\omega})\\
  x_i^{1} & \geq  0 & \quad & \quad (\xi_i)\\
  x_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\varphi_{i,\omega,t})\\
{\color{blue} RP_i - \bar{Q}_i - \bar{Q}_{i,t} - x^{1}_i - \sum_{t > 1} x^{t}_{i,w} } & {\color{blue} \geq 0} & {\color{blue} \forall \quad i,\omega }&  {\color{blue} \quad (\psi_{i,\omega}) }
  \end{align}\\

 (\textbf{FELIPE}: Note that $C_i\seba{(Q_i)}$ represents the \seba{marginal} cost function of producer $i$. We consider quadratic \seba{marginal} cost functions of the form $C_i\seba{(Q_i)}=a+\frac{b}{2}Q\seba{_i}$. Hence, I have rewritten the model formulation to account for this change)

The units of each parameter and variable in Equation (\ref{prodFO}) must be suitable such that the objective function has units of cost (us\$). In Equations (\ref{const.1}) and (\ref{const.2}) the factor $CF_i \cdot \tau$ is added to take into account the real production of electricity.\\

\underline{Auctioneer's problem:}

\begin{align}
    \min_{\theta} & -\theta \pi^{a} \\
    \textrm{s.t \ } &  \phi^{-1}(\epsilon) \sigma^2 + \mu - \theta  \geq 0\\
    &  \theta  \geq 0
\end{align}
\vspace{0.5cm}

\underline{Market clearing constraints:}

\begin{align}
\textrm{(available allowances)}: &  \ \   \sum_{i} A_{i} = \theta   &  & \ \  (\pi^{a})\\
\textrm{(equilibrium in trading market)}: &   \ \  \sum_{i, j\neq i} P_{j,i,\omega} = \sum_{i} V_{i,\omega} & \forall \ \omega & \ \ (\pi^{v}) \\
\textrm{(fulfillment of the demand --first stage)}:  &   \ \  \sum_{i} Q^{1}_i = D^{1}, &  & \ \ (\pi^{d,1})\\
\textrm{(fulfillment of the demand --second stage)}:  &   \ \  \sum_{i} Q^{t}_{i,\omega} = D^{t}_{\omega}, & \forall \ \omega, \tau & \ \ (\pi^{d,t}_\omega)
\end{align}

\vspace{0.5cm}

The Lagrangian function for the producer's problem of each $i \in \{ 1,...,n\}$ is:

\begin{align}
    \mathcal{L}_i(x,Q,A,P,V) = &  {\color{red}(a_iQ_{1,i}+\frac{b}{2}Q_{1,i}^{2})-\pi^{d,1} Q_{1,i}}  + A_i \pi^{a} + I_i x_i^{1}  + \sum_{\Omega} prob_\omega  \Bigg( \sum_{t>1} \frac{1}{(1+R)^t} \Big[ (TC_{i}\cdot {\color{red}(a_iQ_{i,t,\omega}+\frac{b}{2}Q_{i,t,\omega}^{2})}\nonumber \\
    & {\color{red}-\pi^{d,2}_{\omega}) Q_{i,t,\omega}} \nonumber \\
    & + \sum_{t > 1} TCR_i \cdot I_ix^{t}_{i,\omega} \Big]  - V_{i,\omega}\pi_{i,\omega}^v+\sum_{j\neq i}  P_{i,j,\omega}\pi_{j,\omega}^v\Bigg) + \kappa_{i}\Big[Q_{i}^{1} -  \Big(CF_i\cdot\tau \Big)\bar{Q_i}\Big] \nonumber \\ 
   &  +\sum_{\omega,t>1} \alpha_{i,\omega,t}\Bigg[Q_{i,\omega}^{t} - \Big(CF_i \cdot\tau\Big) \Big(\bar{Q_i} +\bar{Q}_{i,t}+ \sum_{t^{\prime}<= t } x^{t}_{i,w} + x_i^{1} \Big)\Bigg]  + \sum_{\omega}\beta_{i,\omega}\Big[V_{i,\omega}-A_i\Big]  \nonumber \\
   & + \sum_{\omega}\gamma_{i,\omega} \Big[-A_{i} - \sum_{j \neq i}P_{i,j,\omega} + V_{i,\omega}+\sum_{t>1}Q_{i, \omega}^{t}\varepsilon_{i}+Q_i^{1}\varepsilon_{i}\Big] \nonumber\\
     & - \sum_{\omega, t>1}\delta_{i,\omega,t} Q_{i,\omega}^{t} - \lambda_{i}\Big[Q^{1}_{i}\Big] - \sum_{\omega, t>1}\varphi_{i,\omega,t} x_{i,\omega}^{t} - \xi_i x^{1}_i {\color{blue} + \sum_{\omega}\psi_{i,\omega} \Big[  \bar{Q}_i+\bar{Q}_{i,t} + x^{1}_i + \sum_{t > 1} x^{t}_{i,w} - RP_i \Big]}
\end{align}

where $\alpha_{i,\omega,\tau}$, $\kappa_i$, $\beta_{i,\omega}$, $\gamma_{i,\omega}$ and $\delta_{i,\omega}$  $\lambda_i$, $\varphi_{i,\omega,t}$ ad $\xi_i$ are the lagrange multipliers of the constraints. \newline

The KKT conditions of the producer's problem are then

\begin{align}
    I_i {\color{blue} + \sum_{\omega}\psi_{i,\omega}} -\sum_{\omega,\tau>1} \alpha_{i,\omega,\tau} -\xi_i & = 0 & \forall i {\color{blue}}&  \qquad (x^{1}_{i})\\
    prob_\omega \Bigg[\frac{1}{(1+R)^t}TCR_i \cdot I_i -  \varphi_{i,\omega,t}\Bigg] - \sum_{t> t\prime}\alpha_{i,\omega,t} ( CF_i \cdot \tau) {\color{blue} + \psi_{i,\omega}}& = 0 & \forall i, \omega, t> 1  &  \qquad (x^{t}_{i,\omega})\\
    {\color{red}(a_{i}+b_iQ^{1}_{i})}-(\pi^{d,1}) + \kappa_i - \lambda_{i} + \sum_{\omega} \gamma_{i,\omega}\varepsilon_i & = 0 & \forall i  &  \qquad (Q^{1}_{i})\\
    prob_\omega \Bigg( \frac{1}{(1+R)^t}\Bigg) ( {\color{red}(TC_i \cdot a_{i}+b_iQ_{i,\omega}^{\tau})}-\pi^{d,t}_{\omega}) + \alpha_{i,\omega,\tau} + \gamma_{i,\omega} \varepsilon_{i}-\delta_{i,\omega,\tau} & =0 & \forall i, \omega, \tau>1 &  \qquad (Q_{i,\omega}^{\tau})\\
    \pi^{a} - \sum_{\omega}\beta_{i,\omega} - \sum_{\omega}\gamma_{i,\omega} & =0 & \forall i & \qquad (A_{i}) \\
    -prob_\omega \pi^v_{i,\omega} + \beta_{i,\omega}  + \gamma_{i,\omega} & =0 & \forall i, \omega & \qquad (V_{i,\omega}) \\
    prob_\omega \pi^v_{j,\omega} -\gamma_{i,\omega} & = 0 & \forall i, j \neq i ,\omega & \qquad (P_{i,j,\omega})
\end{align}

\smallskip

\begin{flushleft}
Primal feasibility
\end{flushleft}

\begin{align}
\big(CF_i \cdot \tau \big) \Bigg[\bar{Q_i} + \sum_{t\leq t^{\prime}} x^{t}_{i,w} + x_i^{1} \Bigg] - Q_{i,\omega}^{t} & \geq 0  & \forall  \quad \omega, t  > 1 & \quad (\alpha_{i,\omega,t}) \\
    \Big(CF_i\cdot\tau \Big)\bar{Q_i}-Q_{i}^{t} & \geq 0  &  \quad t  =1 & \quad (\kappa_i) \\
 A_{i} -V_{i,\omega} & \geq  0  & \forall  \quad \omega & \quad (\beta_{i,\omega})\\
 A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{t>1}Q_{i,\omega}^{t}\varepsilon_{i}-Q_i^{1}\varepsilon_{i} & \geq  0  &\forall \quad \omega & \quad (\gamma_{i,\omega})\\
 Q_i^{1} & \geq  0 & \quad & \quad (\lambda_i)\\
 Q_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\delta_{i,\omega})\\
  x_i^{1} & \geq  0 & \quad & \quad (\xi_i)\\
  x_{i, \omega}^{t} & \geq  0   & \forall  \quad \omega, t >1 & \quad (\varphi_{i,\omega,t})\\
  {\color{blue} RP_i - \bar{Q}_i - x^{1}_i - \sum_{t>1} x^{t}_{i,w} } & {\color{blue} \geq 0} & {\color{blue} \forall \quad i,\omega }&  {\color{blue} \quad (\psi_{i,\omega}) }
\end{align}

\smallskip

\begin{flushleft}
Complementary slackness
\end{flushleft}

\begin{align}
    \Bigg( \big(CF_i \cdot \tau \big) \Bigg[\bar{Q_i} + \sum_{t\leq t^{\prime}} x^{t}_{i,w} + x_i^{1} \Bigg] - Q_{i,\omega}^{t} \Bigg) \cdot \alpha_{i,\omega,\tau} & = 0 & \forall  \quad \omega, t  > 1\\
    \Bigg( \Big(CF_i\cdot\tau \Big)\bar{Q_i}-Q_{i}^{t} \Bigg)\cdot \kappa_i & = 0  & \quad t  =1 \\
    \Bigg(  A_{i} -V_{i,\omega} \Bigg) \cdot \beta_{i,\omega} & = 0 & \forall  \quad \omega\\
    \Bigg( A_{i} + \sum_{j \neq i}P_{i,j,\omega} - V_{i,\omega}-\sum_{t>1}Q_{i,\omega}^{t}\varepsilon_{i}-Q_i^{1}\varepsilon_{i}\Bigg)\cdot \gamma_{i,\omega} & = 0 & \forall \quad \omega \\
    \Bigg( Q_i^{1} \Bigg) \cdot \lambda_i & = 0 & \\
    \Bigg(Q_{i,\omega}^{\tau}\Bigg ) \cdot \delta_{i,\omega} & = 0 & \forall  \quad \omega, t >1\\
    \Bigg( x_i^{1} \Bigg) \cdot \xi_i & = 0 & \\
    \Bigg( x_{i, \omega}^{t} \Bigg) \cdot \varphi_{i,\omega,t} & = 0 & \forall  \quad \omega, t >1 \\
    {\color{blue} \Bigg(RP_i - \bar{Q}_i - x^{1}_i - \sum_{t>1} x^{t}_{i,w}  \Bigg) \cdot \psi_{i,\omega}}&  {\color{blue} = 0 }&  {\color{blue} \forall \quad i,\omega }
RP\end{align}

\smallskip

\begin{flushleft}
Dual Feasibility
\end{flushleft}

\begin{align}
    \alpha_{i,\omega,\tau} & \geq 0 \\
    \kappa_i & \geq 0 \\
    \beta_{i,\omega} & \geq 0 \\
    \gamma_{i,\omega} & \geq 0 \\
    \lambda_i & \geq 0 \\
    \delta_{i,\omega,t} & \geq 0 \\
    \xi_i & \geq 0 \\
    \varphi_{i,\omega,t} & \geq 0 \\
    {\color{blue}\psi_{i,\omega}} & {\color{blue} \geq 0}
\end{align}

For the Auctioneer, the Lagrangian function is

\begin{equation}
    \mathcal{L}(\theta)= -\theta \pi^{a} + \eta (\phi^{-1}(\epsilon) \sigma^2 + \mu - \theta) - \zeta \theta
\end{equation}

where $\eta$ and $\zeta$ are the Lagrange multipliers. Thus, we can obtain the KKT conditions for the auctioneer problem.

\begin{align}
    -\pi^{a} +\eta -\zeta & = 0 \\
    \theta & \geq  0 \\
    (\phi^{-1}(R) \sigma^2 + \mu - \theta) \eta & =  0\\
    \eta & \geq  0\\
    \zeta & \geq 0
\end{align}

\section{Chile's commitment in the Paris Agreement\footnote{https://climateactiontracker.org/countries/chile/2019-06-17/current-policy-projections/ \\https://mma.gob.cl/wp-content/uploads/2015/09/hojaderuta.pdf} }

\begin{itemize}
    \item By 2030: 30\% below 2007 GHG emission (uncondicional target)
    \item By 2035: 60\% electricity production from renewable energy 
    \item By 2050: 70\% electricity production from renewable energy
    \item By 2024: close eight of the oldest coal-fired power plants (20\% of current coal electricity capacity) 
    \item By 2040: phase-out coal 
\end{itemize}

\end{document}